---
title: "Lab 6 - Matt_McKnight"
author: "Matt McKnight"
date: "March 2, 2017"
output:
  html_document:
    runtime: shiny
---

# Graphics
```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )


library( Lahman )
library(tibble)
library(dplyr)
data( Teams )
ave.so <- Teams$SO / Teams$G


```

# Part I
## QUESTION 1 & 2:
### Plot average strike-outs by year. Use the gray() function to select a color, and add transparency. Use appropriate x and y axis conventions and labels.
```{r}
# scatterplot
plot(
      Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
      xlab='', ylab='',
      pch=19,
      col=gray(.7, alpha=.5),
      cex=.65,
      abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
      xlim=c(1900, 2015), ylim=c(0,9.475),
      axes=FALSE, frame.plot=FALSE 
      )

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)


# Used:
# http://www.statmethods.net/advgraphs/axes.html
# http://stackoverflow.com/questions/16702137/remove-y-label-from-plot-in-r
# Many other places for one or two args
```


## QUESTION 3:
### Calculate the league average using tapply() or group_by() year, and add it to the graph. Use the argument type="b" to connect the dots.
```{r}
league_ave <- tapply(ave.so, Teams$yearID, mean)
years<- c(1871:2015)

# Scatterplot
plot(
      Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
      xlab='', ylab='',
      pch=19,
      col=gray(.7, alpha=.5),
      cex=.65,
      abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
      xlim=c(1900, 2015), ylim=c(0,9.475),
      axes=FALSE, frame.plot=FALSE 
      )

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)
     
# Plot league average
points(
       years[years>1899],league_ave[years>1899],
       col="darkslateblue", type="b", pch=19, lwd=1)
points(
       years[years>1899],league_ave[years>1899],
       col="darkslateblue", type="l", pch=19, lwd=1)


```


## QUESTION 4:
### Reproduce at least two of the narrative texts ("US enters World War I", etc.). Note that a line break within text is created by including "\n" in your string.
```{r eval=FALSE}
print("Don't run me")
identify(years,league_ave)
# Note: identified two points for narrative text events: 76 93

```

```{r}

# Scatterplot
plot(
      Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
      xlab='', ylab='',
      pch=19,
      col=gray(.7, alpha=.5),
      cex=.65,
      abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
      xlim=c(1900, 2015), ylim=c(0,9.475),
      axes=FALSE, frame.plot=FALSE 
      )

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)
     
# Plot league average
points(
       years[years>1899],league_ave[years>1899],
       col="darkslateblue", type="b", pch=19, lwd=1)
points(
       years[years>1899],league_ave[years>1899],
       col="darkslateblue", type="l", pch=19, lwd=1)

# Add Text 
text(
     x=years[c(76)], y=league_ave[c(76)], 
     labels= "Players return \n from World War II", 
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=3.5 ) 
text(
     x=years[c(93)], y=league_ave[c(93)],
     labels= "Strike zone enlarged \n from 1963-68",
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=4.5 ) 

segments(
         1946, 3.5, 1946, 1.7, col=gray(.6, alpha=1))
segments(
         1963, 5.5, 1963, 3, col=gray(.6, alpha=1))

mtext(
      "Strikeouts on the Rise ", 
      side=3, line=2, adj=.01, cex=1.5, col="black")
mtext(
      "There were more strikeouts in 2012 than at any other time in major league history.", 
      side=3, line=1, adj=.01, cex=1, col="black")

```



# Part II
## QUESTION 1:
### Using R Shiny tools, add an input panel that allows users to select a team. Once selected, add the team trendline to the graph.
```{r eval=FALSE}
print("Don't run me")

library(shiny)

selectInput(inputId="team", 
            label="Strikeouts per game per team (by batters)", 
            choices= unique(Teams$name),
            selected="Chicago Cubs"
            )

renderPlot({

attach( Teams )

# Scatterplot
plot(
      Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
      xlab='', ylab='',
      pch=19,
      col=gray(.7, alpha=.5),
      cex=.65,
      abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
      xlim=c(1900, 2015), ylim=c(0,9.475),
      axes=FALSE, frame.plot=FALSE 
      )

# Plot league average
points(
       names(league_ave), league_ave, type="b", 
       col="darkslateblue", pch=19, cex=1)
points(
       names(league_ave), league_ave, type="l", 
       col="darkslateblue", pch=19, cex=1)

# Add legend
legend(
       1899, 8.7, "League Average", bty="n", 
       lty=1, lwd=2.5, col="darkslateblue")

# Plot user input team average 
user.team <- filter( Teams , name == input$team )
user.team.so <- user.team$SO / user.team$G
points( 
       user.team$yearID, user.team.so,  type="b", 
       lwd = 2 , pch = 19 , col="darkgoldenrod1" )
points( 
       user.team$yearID, user.team.so,  type="l", 
       lwd = 2 , pch = 19 , col="darkgoldenrod1" )

# Add legend
legend(
       1899, 8.1, "Team Average", bty="n", 
       lty=1, lwd=2.5, col="goldenrod1")

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)

# Add text 
text(
     x=years[c(76)], y=league_ave[c(76)], 
     labels= "Players return \n from World War II", 
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=3.5 ) 
text(
     x=years[c(93)], y=league_ave[c(93)],
     labels= "Strike zone enlarged \n from 1963-68",
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=4.5 ) 

segments(
         1946, 3.5, 1946, 1.7, col=gray(.6, alpha=1))
segments(
         1963, 5.5, 1963, 3, col=gray(.6, alpha=1))

mtext(
      "Strikeouts on the Rise ", 
      side=3, line=2, adj=.01, cex=1.5, col="black")
mtext(
      "There were more strikeouts in 2012 than at any other time in major league history.", 
      side=3, line=1, adj=.01, cex=1, col="black")

})

```

```{r, echo=FALSE} 
library(shiny)

selectInput( inputId="team", 
             label="Strikeouts per game per team (by batters)", 
             choices= unique(Teams$name),
             selected="Chicago Cubs"
           )

renderPlot({

attach( Teams )

# Scatterplot
plot(
      Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
      xlab='', ylab='',
      pch=19,
      col=gray(.7, alpha=.5),
      cex=.65,
      abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
      xlim=c(1900, 2015), ylim=c(0,9.475),
      axes=FALSE, frame.plot=FALSE 
      )

# Plot league average
points(
       names(league_ave), league_ave, type="b", 
       col="darkslateblue", pch=19, cex=1)
points(
       names(league_ave), league_ave, type="l", 
       col="darkslateblue", pch=19, cex=1)

# Add legend
legend(
       1899, 8.7, "League Average", bty="n", 
       lty=1, lwd=2.5, col="darkslateblue")

# Plot user input team average 
user.team <- filter( Teams , name == input$team )
user.team.so <- user.team$SO / user.team$G
points( 
       user.team$yearID, user.team.so,  type="b", 
       lwd = 2 , pch = 19 , col="darkgoldenrod1" )
points( 
       user.team$yearID, user.team.so,  type="l", 
       lwd = 2 , pch = 19 , col="darkgoldenrod1" )

# Add legend
legend(
       1899, 8.1, "Team Average", bty="n", 
       lty=1, lwd=2.5, col="goldenrod1")

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)

# Add text 
text(
     x=years[c(76)], y=league_ave[c(76)], 
     labels= "Players return \n from World War II", 
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=3.5 ) 
text(
     x=years[c(93)], y=league_ave[c(93)],
     labels= "Strike zone enlarged \n from 1963-68",
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=4.5 ) 

segments(
         1946, 3.5, 1946, 1.7, col=gray(.6, alpha=1))
segments(
         1963, 5.5, 1963, 3, col=gray(.6, alpha=1))

mtext(
      "Strikeouts on the Rise ", 
      side=3, line=2, adj=.01, cex=1.5, col="black")
mtext(
      "There were more strikeouts in 2012 than at any other time in major league history.", 
      side=3, line=1, adj=.01, cex=1, col="black")

})
```


##QUESTION 2:
###Report the selected team name on the graph.
```{r}
library(shiny)
library( Lahman )
data(Teams)
ave.so <- Teams$SO / Teams$G
league_ave <- tapply(ave.so, Teams$yearID, mean)
years<- c(1871:2015)

#selection code adapted from https://shiny.rstudio.com/articles/plot-interaction.html

ui <- basicPage(
plotOutput("plot1",
           hover = "plot_hover",
           click = "plot_click"
           ),
           verbatimTextOutput("info"))

server <- function(input, output) {
  output$plot1 <- renderPlot({
    plot(
          Teams$yearID[Teams$yearID>1899], ave.so[Teams$yearID>1899], 
          xlab='', ylab='',
          pch=19,
          col=gray(.7, alpha=.5),
          cex=.65,
          abline(h=1:9, col = gray(.7, alpha=.5), lty =3),
          xlim=c(1900, 2015), ylim=c(0,9.475),
          axes=FALSE, frame.plot=FALSE 
         )

# Plot league average
points(
       names(league_ave), league_ave, type="b", 
       col="darkslateblue", pch=19, cex=1)
points(
       names(league_ave), league_ave, type="l", 
       col="darkslateblue", pch=19, cex=1)

# Add legend
legend(
       1899, 8.7, "League Average", bty="n", 
       lty=1, lwd=2.5, col="darkslateblue")

# Add legend
legend(
       1899, 8.1, "Team Average", bty="n", 
       lty=1, lwd=2.5, col="goldenrod1")

# Add axes
axis( 
     side=1, at=c(1900,2015), labels=c("",""), lwd.ticks=0)
axis( 
     side=1, at=seq (1900, 2020, by=10 ), 
     padj=-2, hadj=-.08, tck=-.015, cex.axis=0.8)

axis( 
     side=4, at=seq(0, 9, by=1 ), las=2,  
     lty=0, col.axis=gray(.4, alpha=1),
     tick=FALSE, line=-.9, tck=0, cex.axis=0.8)

# Add text 
text(
     x=years[c(76)], y=league_ave[c(76)], 
     labels= "Players return \n from World War II", 
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=3.5 ) 
text(
     x=years[c(93)], y=league_ave[c(93)],
     labels= "Strike zone enlarged \n from 1963-68",
     pos=1, cex=.6, col=gray(.6, alpha=1), offset=4.5 ) 

segments(
         1946, 3.5, 1946, 1.7, col=gray(.6, alpha=1))
segments(
         1963, 5.5, 1963, 3, col=gray(.6, alpha=1))

mtext(
      "Strikeouts on the Rise ", 
      side=3, line=2, adj=.01, cex=1.5, col="black")
mtext(
      "There were more strikeouts in 2012 than at any other time in major league history.", 
      side=3, line=1, adj=.01, cex=1, col="black")
  })

# Create click and hover tools
  output$info <- renderText({
    xy_str <- function(e) {
      if(is.null(e)) return("NULL\n")
      paste0("x=", round(e$x, 1), " y=", round(e$y, 1), "\n")
    }
    xy_range_str <- function(e) {
      if(is.null(e)) return("NULL\n")
      paste0("xmin=", round(e$xmin, 1), " xmax=", round(e$xmax, 1), 
             " ymin=", round(e$ymin, 1), " ymax=", round(e$ymax, 1))
    }

    paste0(
      "click: ", xy_str(input$plot_click),
      "hover: ", xy_str(input$plot_hover)
    )
  })
 
}

shinyApp(ui, server)
```
Note: I was unsure how to combine the click/hover shiny tool with the select drop-down menu shiny tool, so I ran both separately.

