---
title: "Lab 2"
author: "Matt McKnight"
date: "February 3rd, 2017"
output:
  word_document: default
  html_document: default
---

## Syracuse Land Parcel Database
```{r}
dat <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv")
ls(dat)

```



## Question 1
### What is the total assessed value of all the land in Syracuse?
```{r}
sum(dat$AssessedVa)
```
## Answer
###The total assessed value of all the land in Syracuse is $7,538,424,950.





## Question 2
### What proportion of land parcels is tax exempt? Use the LandUse variable and the table below to create a logical statement that identified all of the tax exempt parcels. What is the value of these non-taxable parcels?
```{r}
#Basically create a logical vector. 

# Creation of land parcels tax exempt dichotomous variable.
NonTaxed_land <- c(dat$LandUse == "Vacant Land" | dat$LandUse == "Parking" | dat$LandUse == "Schools" | dat$LandUse == "Parks" | dat$LandUse == "Cemetery" | dat$LandUse == "Religious" | dat$LandUse == "Recreation" | dat$LandUse == "Community Services" | dat$LandUse == "Utilities")

# Number of land parcels tax exempt. 
sum(dat$LandUse == "Vacant Land" | dat$LandUse == "Parking" | dat$LandUse == "Schools" | dat$LandUse == "Parks" | dat$LandUse == "Cemetery" | dat$LandUse == "Religious" | dat$LandUse == "Recreation" | dat$LandUse == "Community Services" | dat$LandUse == "Utilities")

#Find values for Taxed land parcels, FALSE category
summary(NonTaxed_land)

#Porportion of land parcels Tax exempt
4878/(4878+36624)

#Value of non-taxable parcels
sum(dat$AssessedVa[NonTaxed_land == TRUE])

```
## Answer

Twelve percent of land parcels in Syracuse are tax exempt. The value of these non-taxable land parcels is $3,006,860,002.



## Question 3
### Which property has the highest delinquent tax bill?
```{r}
which.max(dat$AmtDelinqu) 
max(dat$AmtDelinqu)

```
## Answer
### The property with the highest delinquent tax bill was N #22469 with $1,543,354. Further review of #22469 in the data spreadsheet shows that this is a commercial property downtown.



## Question 4
### Which of these neighborhoods listed below has the highest proportion of vacant buildings?
```{r}
table(dat$Nhood, dat$VacantBuil)
# Westcott porportion
24/(1474+24) 
# Northside porportion
264/(2977+264) 
#Strathmore proportion
38/(1769+38) 
#Southside proportion
120/(1239+120) 
0.088*100
```
## Answer

The Southside neighborhood reports the highest porportion of vacant buildings at 8.8 percent.



## Question 5
### How many single family homes worth more than $100,000 each are located in each of the four neighborhoods above?
```{r}

#True or False Statements
#dat$LandUse == "Single Family" | dat$AssessedVa > 100000 | dat$Nhood == "Westcott" | dat$Nhood == "Northside" | dat$Nhood == "Strathmore" | dat$Nhood == "Southside"

#Single family homes > $100,000 in Westcott
sum(dat$LandUse == "Single Family" & dat$AssessedVa > 100000 & dat$Nhood == "Westcott")

#Single family homes > $100,000 in Northside
sum(dat$LandUse == "Single Family" & dat$AssessedVa > 100000 & dat$Nhood == "Northside")

#Single family homes > $100,000 in Strathmore
sum(dat$LandUse == "Single Family" & dat$AssessedVa > 100000 & dat$Nhood == "Strathmore")
     
#Single family homes > $100,000 in Southside
sum(dat$LandUse == "Single Family" & dat$AssessedVa > 100000 & dat$Nhood == "Southside")

#Sum of all select neighborhoods
208+5+461+0

#Should learn how to place all of these responses in a table to present!
```
## Answer

The number of single family homes over $100,000 in four Syrcause neighborhoods are as follows:

Westcott   208  
Northside  5  
Strathmore 461

Southside  0 

In total, these select single family homes number 674 in the four selected neighborhoods.



