---
title: "Lab 1"
author: "Matt McKnight"
date: "January 28, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

### Beginning command Lines
```{r}
library("Lahman")
data("Master")
ls(Master)
```

The command "ls" lists the variables found within the Master dataset.

## Question 1
### What is the average weight of an MLB player?
```{r}
summary(Master$weight)
mean(Master$weight, na.rm = TRUE)
```

The average weight of a MLB player is 186 pounds.

## Question 2
### How many MLB players have come from the Netherlands? 
```{r}
table(Master$birthCountry)
table(Master$birthCountry == "Netherlands")
```

The number of MLB players born in the Netherlands is 12.

## Question 3
### How tall was the tallest player in MLB? The shortest?
```{r}
summary(Master$height)
```

The tallest MLB player was 6'11" inches, while the shortest was only 3'7" inches.

## Question 4
### What percentage of players switch hit (bat from both sides of the plate)? 
```{r}
table(Master$bats)
(1163/(1163+4879+11614))*100
```

6.59 percent of MLB players are switch hitters.

## Question 5
### OPTIONAL: How many players are right-handed when they throw, but left-handed when they bat?
```{r}
table(Master$throws, Master$bats)
sum(Master$throws == "R" & Master$bats == "L", na.rm = TRUE)
```

2,054 players throw right-handed but bat left-handed.

