---
title: "Lab 5 - Matt_McKnight"
author: "Matt McKnight"
date: "February 25, 2017"
output:
  pdf_document: default
  html_document:
    df_print: paged
    keep_md: yes
---

# Plot Function

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )

library(Lahman)
data(Teams)

```

## QUESTION 1
###	Drop all of the data before 1990.
```{r}
which(Teams$yearID == 1990)
Teams_90 <- Teams[-c(1:2047), ]

```


## QUESTION 2
### Examine the relationship between Hits and Wins using a scatterplot. Use meaningful labels and a visually appealing style.
```{r}
plot(
      x=Teams_90$H, y=Teams_90$W, 
      main="Key Performance Indicators: Offense",
      xlab="Total Hits Per Season Per Team", ylab="Total Wins",
      pch=19,
      col="darkorchid1",
      cex=.75,
      bty="L"
      )


```



## QUESTION 3
### Annotate two outliers on the graph with the teamID. 
```{r}
#identify ( Teams_90$H, Teams_90$W )

plot(
      x=Teams_90$H, y=Teams_90$W, 
      main="Key Performance Indicators: Offense",
      xlab="Total Hits Per Season Per Team", ylab="Total Wins",
      pch=19,
      col="darkorchid1",
      cex=.75,
      bty="L",

      text( x=Teams_90$H[c(333,379)], y=Teams_90$W[c(333,379)],
      labels=Teams_90$teamID[c(333,379)], pos=4 )
      )


```



## QUESTION 4
### Examine the relationship between ERA and Wins using a scatterplot. Use meaningful labels and a visually appealing style.
```{r}
plot(
      x=Teams_90$ERA, y=Teams_90$W, 
      main="Key Performance Indicators: Offense",
      xlab="Earned Run Average Per Season Per Team", ylab="Total Wins",
      pch=19,
      col="paleturquoise3",
      cex=.75,
      bty="L"
      )


```



## QUESTION 5
### Annotate two outliers on the graph with the teamID.
```{r}
#identify ( Teams_90$ERA, Teams_90$W )

plot(
      x=Teams_90$ERA, y=Teams_90$W, 
      main="Key Performance Indicators: Offense",
      xlab="Earned Run Average Per Season Per Team", ylab="Total Wins",
      pch=19,
      col="paleturquoise3",
      cex=.75,
      bty="L",

      text( x=Teams_90$ERA[c(333,379)], y=Teams_90$W[c(333,379)],
      labels=Teams_90$teamID[c(333,379)], pos=2 )
      )


```


## QUESTION 6 - BONUS
### Add a trend line to the scatterplot to highlight the relationship.
```{r}
plot(
      x=Teams_90$ERA, y=Teams_90$W, 
      main="Key Performance Indicators: Offense",
      xlab="Earned Run Average Per Season Per Team", ylab="Total Wins",
      pch=19,
      col="paleturquoise3",
      cex=.75,
      bty="L",

      text( x=Teams_90$ERA[c(333,379)], y=Teams_90$W[c(333,379)],
      labels=Teams_90$teamID[c(333,379)], pos=2 )
      )

lines(lowess(Teams_90$ERA,Teams_90$W), col="darkblue", lwd=3)

```





